---
title: "p8105_hw6_cl4776"
author: "Chenxi Liu"
date: "2025-12-01"
output: github_document
---
# Problem 1 
```{r,knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)}

library(tidyverse)
library(broom)
library(purrr)

homicides <- readr::read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")


```

```{r}
homicides_clean <- homicides |>
mutate(
city_state = str_c(city, ", ", state),
resolved   = if_else(disposition == "Closed by arrest", 1, 0),
victim_age = as.numeric(victim_age),
victim_sex = factor(victim_sex, levels = c("Female", "Male")),
victim_race = factor(victim_race)
) |>
filter(
!city_state %in% c("Dallas, TX", "Phoenix, AZ",
"Kansas City, MO", "Tulsa, AL"),
victim_race %in% c("White", "Black")
)

summary(homicides_clean$victim_age)
count(homicides_clean, city_state)
baltimore <- homicides_clean |>
filter(city_state == "Baltimore, MD")

balt_glm <- glm(
resolved ~ victim_age + victim_sex + victim_race,
family = binomial,
data   = baltimore
)

balt_tidy <- balt_glm |>
tidy(conf.int = TRUE, exponentiate = TRUE)

balt_tidy

balt_OR_male <- balt_tidy |>
filter(term == "victim_sexMale") |>
select(term, estimate, conf.low, conf.high)

balt_OR_male
```

```{r}
city_models <- homicides_clean |>
group_by(city_state) |>
nest() |>
mutate(
model = map(
data,
~ glm(
resolved ~ victim_age + victim_sex + victim_race,
family = binomial,
data = .x
)
),
tidied = map(
model,
~ tidy(.x, conf.int = TRUE, exponentiate = TRUE)
)
) |>
unnest(tidied) |>
filter(term == "victim_sexMale") |>
select(city_state, estimate, conf.low, conf.high) |>
ungroup()

city_models
city_OR_plot <- city_models |>
arrange(estimate) |>
mutate(city_state = factor(city_state, levels = city_state)) |>
ggplot(aes(x = city_state, y = estimate)) +
geom_hline(yintercept = 1, linetype = "dashed") +
geom_point() +
geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
coord_flip() +
labs(
title = "Adjusted Odds Ratios for Solving Homicides\nMale vs Female Victims",
x = "City",
y = "Odds Ratio (male vs female victims)"
) +
theme_minimal()

city_OR_plot
```

The plot displays the adjusted odds ratio (OR) for solving homicides comparing male to female victims across major U.S. cities, controlling for victim age and race. Most cities have OR estimates close to 1 with confidence intervals crossing 1, indicating little evidence of a meaningful difference in clearance rates between male and female victims. A few cities show ORs above or below 1, but wide confidence intervals reflect substantial uncertainty or small sample sizes. An odds ratio of 1 indicates no difference in clearance probability between male and female victims, while values above or below 1 represent higher or lower odds of being solved for male victims, respectively. Overall, the results suggest that after adjustment, victim sex is not a strong predictor of whether a homicide is solved in most cities.

# Problem 2
```{r}
library(p8105.datasets)
data("weather_df")
```
```{r}
weather_clean <- weather_df |>
drop_na(tmax, tmin, prcp)
set.seed(1)

bootstrap_results <- tibble(i = 1:5000) |>
mutate(
sample_df = map(i, ~ weather_clean |> slice_sample(prop = 1, replace = TRUE)),
models = map(sample_df, ~ lm(tmax ~ tmin + prcp, data = .x)),
glance_out = map(models, glance),
tidy_out   = map(models, tidy)
) |>
mutate(
r_squared = map_dbl(glance_out, ~ .x$r.squared),
beta_prod = map_dfr(tidy_out, ~
.x |>
filter(term %in% c("tmin", "prcp")) |>
summarise(prod = prod(estimate))
)$prod
)

bootstrap_results |>
ggplot(aes(x = r_squared)) +
geom_histogram(bins = 40, color = "white") +
theme_minimal() +
labs(
title = "Bootstrap Distribution of R²",
x = "R²",
y = "Count"
)
```

```{r}
bootstrap_results |>
ggplot(aes(x = beta_prod)) +
geom_histogram(bins = 40, color = "white") +
theme_minimal() +
labs(
title = "Bootstrap Distribution of β₁ × β₂",
x = "β₁ × β₂ (tmin × prcp)",
y = "Count"
)
r2_CI <- quantile(bootstrap_results$r_squared, probs = c(0.025, 0.975))
beta_prod_CI <- quantile(bootstrap_results$beta_prod, probs = c(0.025, 0.975))

r2_CI
beta_prod_CI
```

The bootstrap distribution for R² is approximately symmetric and centered near the value obtained from the original data, indicating that the fitted model explains a consistent proportion of the variation in tmax across bootstrap samples. The distribution shows moderate variability, and the 95% confidence interval reflects reasonable uncertainty in the proportion of variance explained by tmin and prcp.

The distribution of β₁ × β₂ (the product of the coefficients for tmin and prcp) is more skewed, with most values near zero but a noticeable tail, likely because prcp has a small and variable effect. This skewness suggests that the interaction-like quantity β₁β₂ is less stable across bootstrap samples. The 95% confidence interval captures this skewed uncertainty and includes values close to zero.

# Problem 3
```{r}
library(p8105.datasets)
birthweight <- read_csv("data/birthweight.csv")
```
```{r}
birthweight_clean <- birthweight |>
mutate(
babysex = factor(babysex, levels = c(1, 2), labels = c("male", "female")),
malform = factor(malform, levels = c(0, 1), labels = c("absent", "present")),
mrace = factor(mrace, levels = c(1, 2, 3, 4, 8),
labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9),
labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown"))
)

birthweight_clean |>
summarise(across(everything(), ~ sum(is.na(.))))

library(modelr)

bwt_mod_main <- lm(
bwt ~ gaweeks + babysex + bhead + blength + delwt +
wtgain + smoken + ppbmi + momage + mheight + mrace,
data = birthweight_clean
)

summary(bwt_mod_main)
bwt_aug <- birthweight_clean |>
add_predictions(bwt_mod_main) |>
add_residuals(bwt_mod_main)

bwt_aug |>
ggplot(aes(x = pred, y = resid)) +
geom_point(alpha = 0.4) +
geom_hline(yintercept = 0, linetype = "dashed") +
theme_minimal() +
labs(
title = "Residuals vs Fitted Values for Main Birthweight Model",
x = "Fitted values",
y = "Residuals"
)
```

```{r}
bwt_mod_len_gage <- lm(
bwt ~ blength + gaweeks,
data = birthweight_clean
)

bwt_mod_head_len_sex_int <- lm(
bwt ~ bhead * blength * babysex,
data = birthweight_clean
)
set.seed(1)

cv_df <- crossv_mc(birthweight_clean, 100) |>
mutate(
train = map(train, as_tibble),
test  = map(test,  as_tibble),
mod_main = map(train, ~ lm(
bwt ~ gaweeks + babysex + bhead + blength + delwt +
wtgain + smoken + ppbmi + momage + mheight + mrace,
data = .x
)),
mod_len_gage = map(train, ~ lm(
bwt ~ blength + gaweeks,
data = .x
)),
mod_head_len_sex_int = map(train, ~ lm(
bwt ~ bhead * blength * babysex,
data = .x
))
) |>
mutate(
rmse_main = map2_dbl(mod_main, test, ~ rmse(model = .x, data = .y)),
rmse_len_gage =
map2_dbl(mod_len_gage, test, ~ rmse(model = .x, data = .y)),
rmse_head_len_sex_int =
map2_dbl(mod_head_len_sex_int, test, ~ rmse(model = .x, data = .y))
)

cv_df |>
summarise(
rmse_main = mean(rmse_main),
rmse_len_gage = mean(rmse_len_gage),
rmse_head_len_sex_int = mean(rmse_head_len_sex_int)
)
```

For the birthweight data, I cleaned the dataset by converting categorical variables (babysex, malform, mrace, frace) to factors and renaming mheigth to mheight. The data contained essentially no missing values.

My proposed model includes gestational age, sex, head circumference, length, maternal delivery weight, weight gain, pre-pregnancy BMI, smoking, maternal age, height, and race. This set of predictors is motivated by known biological factors related to fetal growth. The residual–fitted plot shows no major patterns, suggesting the model is reasonably appropriate.

I compared this model to two alternatives: (1) a simple model using only length and gestational age, and (2) a complex interaction model including head circumference, length, sex, and all interactions. Using 100 Monte Carlo cross-validation splits, the proposed model had the lowest average RMSE. The simple model performed worse due to missing important predictors, and the interaction model performed slightly worse due to overfitting. Overall, the proposed model provides the best balance of interpretability and predictive accuracy.
